---
title: "Basic Data Manipulation"
author: "Aren Simmons"
date: "10/31/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Wrangling Exercise 1
Clean this data set to make it easier to visualize and analyze

```{r message = FALSE, warning= FALSE}
library(dplyr)
library(tidyr)
library(readr)
Data <- read_csv("refine - Sheet1.csv")
tbl_df(Data)
```

## 1: Clean up brand names to standardize them and remove misspellings

```{r}
clean_companies <- Data %>%
mutate(company = tolower(company))%>%
mutate(company = 
ifelse((substr(company,1,2) == "ak"), "akzo",
       ifelse((substr(company,1,2) == "ph"|substr(company,1,2) =="fi"),"phillips",
       ifelse((substr(company,1,2) == "un"), "unilever", 
              ifelse((substr(company,1,2) == "va"), "van houten", company)))))
View(clean_companies)
```

## 2: Separate product code and number into two columns for product code and product number

```{r}
separate_product <- separate(clean_companies,2, c("product_code","product_number"))
View(separate_product)
```

## 3: Assigning the product code to represent the product categories
```{r}
product_categories <- separate_product %>% 
  mutate("product_categories" = product_code) %>%
  mutate(product_categories =
           ifelse(product_categories == "p","Smartphone",
         ifelse(product_categories == "v", "TV",
                ifelse(product_categories == "x", "Laptop",
                       ifelse(product_categories == "q", "Tablet", "na")))))
          View(product_categories)
```

## 4: Combining the three address fields (address, city, country) to create a full address
```{r}
full_address <- product_categories %>% 
  unite("full_address",address:country,sep = ",", remove = TRUE)
View(full_address)
```

## 5: Creating dummy variables for company and product category
```{r}
dummy_var <- full_address %>% mutate("company_phillips" = company) %>%
  mutate(company_phillips = ifelse(company_phillips == "phillips", "1",0 ))%>%
  mutate("company_akzo" = company) %>%
  mutate(company_akzo = ifelse(company_akzo == "akzo", "1", 0)) %>%
  mutate("company_van_houten" = company) %>%
  mutate(company_van_houten = ifelse(company_van_houten == "van houten","1", 0))%>%
  mutate("company_unilever" = company) %>%
  mutate(company_unilever = ifelse(company_unilever == "unilever", "1", 0))%>%
  mutate("product_smartphone" = product_categories) %>%
  mutate(product_smartphone = ifelse(product_smartphone == "Smartphone", "1",0))%>%
  mutate("product_tv"= product_categories)%>%
  mutate(product_tv = ifelse(product_tv == "TV","1",0))%>%
  mutate("product_laptop"=product_categories)%>%
  mutate(product_laptop = ifelse(product_laptop == "Laptop", "1", 0))%>%
  mutate("product_tablet" = product_categories)%>%
  mutate(product_tablet = ifelse(product_tablet == "Tablet","1",0))

  View(dummy_var)
```