---
title: "Logistic Regression Mini Project"
author: "SimmonsMo"
date: "1/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise: logistic regression
## ───────────────────────────────────

##   Use the NH11 data set that we loaded earlier.

##   1. Use glm to conduct a logistic regression to predict ever worked
##      (everwrk) using age (age_p) and marital status (r_maritl).
##   2. Predict the probability of working for each level of marital
##      status.

Load the NH11 data
```{r}
library(tidyr)
library(dplyr)
library(magrittr)

NH11 <- readRDS("dataSets/NatHealth2011.rds")
```
Check the structure of NH11
```{r}
str(NH11)
```
Check the structure of Ever worked 
```{r}
str(NH11$everwrk)
levels(NH11$everwrk)
```

Removing missing values from Ever Worked
```{r}
NH11$everwrk <- factor(NH11$everwrk, levels= c("2 No", "1 Yes"))
everwork <- NH11$everwrk
```

Running our regression model
```{r}
EverWorked <- glm(everwork ~age_p + r_maritl, data = NH11, family = "binomial")
coef(summary(EverWorked))
EverWorked.Tab <- coef(summary(EverWorked))
EverWorked.Tab[,"Estimate"] <- exp(coef(EverWorked))
EverWorked.Tab
```
Our model shows us that age, widowed status, divorced, never married and living with a partner are significant variables when it comes to predicting if a patient has ever worked. 

Next we need to create a data frame of the variables that we need for our predictive model.
```{r}
nh_lite <- NH11 %>% 
  select(everwrk, age_p, r_maritl)
```
 We need to check the variables to affirm that they are cleaned for modeling.

Cleaning the NA values in the everwrk column
```{r}
nh_lite2 <- nh_lite %>% 
  filter(!is.na(everwrk))


#head(nh_lite)
#head(nh_lite2)

#str(nh_lite)
#str(nh_lite2)
```

Checking whether our other variables need to be cleaned
```{r}
age_p_distinct <- nh_lite %>% 
  select(age_p) %>%
  distinct %>%
  arrange(age_p)
age_p_distinct
```
Age does not need to be cleaned

Checking levels of marital status
```{r}
r_maritl_distinct <- nh_lite %>%
  select(r_maritl) %>%
  distinct %>%
  arrange(r_maritl)

r_maritl_distinct
```

We want to find the probability of ever working for each level of marital status, therefore nothing needs to be changed. 
```{r}
predData <- with(nh_lite2, 
                expand.grid(everwrk = "1 Yes",
                            r_maritl = nh_lite2
                            %>% select(r_maritl) %>%
                              distinct %>%
                              extract2(1),
                            age_p = mean(age_p, na.rm = TRUE)))
predData

cbind(predData, predict(EverWorked, type = "response",
                       se.fit = TRUE, interval="confidence",
                       newdata = predData))
```
Our results are given above for someone who is 56 years old as a probability in the "fit"" column. 